---
title: "DepMap Portal Data"
author: "gDR team"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{DepMap Portal Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
library(gDRtestData)
```

# Introduction
The gDRtestData package includes a curated subset of DepMap Public 24Q4 data,
serving multiple purposes:

 - test data: validate gDR analysis functions with realistic genomic datasets
 - example data: demonstrate gDR package capabilities in documentation and vignettes
 - reproducible research: enable reproducible examples across the gDR platform.
 
This vignette describes the included DepMap datasets, their contents, and how to use them.

## Data Source & Citation

**Orginal source**: [DepMap Portal](https://depmap.org/portal) 

**Release**: DepMap Public 24Q4 (Loaded May 26, 2026)

**Citation**:
DepMap, Broad (2024). DepMap 24Q4 Public. Figshare+. Dataset. 
https://doi.org/10.25452/figshare.plus.27993248.v1

# Data Overview
The DepMap 24Q4 release contains new cell models and data from:

 - Whole Genome/Exome Sequencing (Copy Number and Mutation)
 - RNA Sequencing (Expression and Fusions)
 - Genome-wide CRISPR knockout screens.

The following datasets are included in the `gDRtestData` pacakge:

| Dataset | Type | Dimensions | Description | Dataset url |
| --- | --- | --- | -------------| --- |
| **Models** | Metadata | ~1,000 cell lines | Cell line information and annotations | [url](https://depmap.org/portal/data_page/?tab=allData&releasename=DepMap%20Public%2024Q4&filename=Model.csv) |
| **CRISPRGeneEffect** | Functional | ~1,000 × ~18,000 | CRISPR knockout gene effect scores (integrated via Chronos) | [url](https://depmap.org/portal/data_page/?tab=allData&releasename=DepMap%20Public%2024Q4&filename=CRISPRGeneEffect.csv) |
| **Expression** | Omics | ~1,000 × ~19,000 | Gene expression (log2 TPM, protein-coding genes) | [url](https://depmap.org/portal/data_page/?tab=allData&releasename=DepMap%20Public%2024Q4&filename=OmicsExpressionProteinCodingGenesTPMLogp1.csv) |
| **Mutations (Hotspot)** | Somatic | ~1,000 × ~3,000 | Binary matrix of hotspot mutations | [url](https://depmap.org/portal/data_page/?tab=allData&releasename=DepMap%20Public%2024Q4&filename=OmicsSomaticMutationsMatrixHotspot.csv) |
| **Mutations (Damaging)** | Somatic | ~1,000 × ~3,000 | Binary matrix of damaging mutations | [url](https://depmap.org/portal/data_page/?tab=allData&releasename=DepMap%20Public%2024Q4&filename=OmicsSomaticMutationsMatrixDamaging.csv) |
| **OmicsCNGene** | CNV | ~1,000 × ~20,000 | Gene-level copy number estimates | [url](https://depmap.org/portal/data_page/?tab=allData&releasename=DepMap%20Public%2024Q4&filename=OmicsCNGene.csv) |

## Data Dictionary

### Models (Cell Line Metadata)

**file name** Model.csv

| Aspect | Details |
|--------|---------|
| **Rows** | Individual cell lines |
| **Columns** | Metadata columns (see below) |
| **Values** | Cell line annotations and metadata |
| **Data Type** | Mixed (character, numeric) |
| **Interpretation** | Comprehensive metadata for each cell line model |
| **NA Handling** | Missing values indicate information not available for that model |


**Column Details:**

Column Summary:

- `ModelID`: Unique cell line identifier
- `CCLEName`: Cell line name from CCLE database
- `CellLineName`: Common cell line name
- `TissueOrigin`: Tissue type (Human, Mouse, Other)
- `DepmapModelType`, `OncotreeLineage`, `OncotreePrimaryDisease`, `OncotreeSubtype`: Cancer classification (from Oncotree)
- `OncotreeCode`: Oncotree classification code
- `PrimaryOrMetastasis`: Tumor site (Primary/Metastatic/Recurrence)
- `Age`: Age at sampling 
- `AgeCategory`: Age category at time of sampling (Adult/Pediatric/Fetus/Unknown)
- `Sex`: Sex at sampling (Female/Male/Unknown)
- `PatientRace`: Patient-reported race

Notes:

- Classification: Oncotree taxonomy for cancer models
- Quality: Authenticated, high-quality cell lines only
- Completeness: Some fields may be NA; indicates information not available
- Use: Primary reference for cell line metadata; join with other datasets via ModelID

### Somatic Mutations (Hotspot)

**file name**: OmicsSomaticMutationsMatrixHotspot.csv

| Aspect | Details |
|--------|---------|
| **Rows** | Cell line identifiers |
| **Columns** | NCBI gene IDs |
| **Values** | Binary (0/1); 1 = hotspot mutation present |
| **Definition** | Mutations in known cancer hotspots (COSMIC, OncoKB) |
| **Sequencing** | Whole exome sequencing (WES) |

Note: Recurrent mutations at known oncogenic positions.

### Somatic Mutations (Damaging)

**file name**: OmicsSomaticMutationsMatrixDamaging.csv

| Aspect | Details |
|--------|---------|
| **Rows** | Cell line identifiers |
| **Columns** | NCBI gene IDs |
| **Values** | Binary (0/1); 1 = damaging mutation present |
| **Definition** | Frame-shift, stop-gain, or splice-site mutations |
| **Quality** | High confidence damaging variants |

Note: Loss-of-function mutations (frameshifts, nonsense, etc.)

### CRISPR Gene Effect

**file name**: CRISPRGeneEffect.csv

| Aspect | Details |
|--------|---------|
| **Rows** | Cell line identifiers |
| **Columns** | NCBI gene IDs (Entrez format) as column names |
| **Values** | CRISPR knockout effect scores |
| **Scale** | -1 to +1 (typically); negative = essential gene in that cell line |
| **Interpretation** | Lower values indicate genes more essential for cell viability |
| **NA Handling** | Missing values indicate insufficient screen coverage |

Note:

- Method: Genome-wide CRISPR/Cas9 knockout screens
- Scale: Dependency scores (probability of essentiality)
- Processing: Already normalized and quality-filtered by DepMap

### Gene Expression

**file name**: OmicsExpressionProteinCodingGenesTPMLogp1.csv

| Aspect | Details |
|--------|---------|
| **Rows** | Cell line identifiers |
| **Columns** | NCBI gene IDs (protein-coding genes only) |
| **Values** | Expression levels (numeric) |
| **Scale** | Log2(TPM + 1); already log-transformed |
| **Range** | Typically 0-20 (log2 scale) |
| **Quality** | RNA-seq from standardized Broad CCL protocols |

Note:

- Only protein-coding genes included
- Already log-transformed (TPM + 1 pseudocount)
- Row-wise and gene-wise normalization already applied by DepMap

### Copy Number Variation (CNV)

**file name**: OmicsCNGene.csv

| Aspect | Details |
|--------|---------|
| **Rows** | Cell line identifiers |
| **Columns** | NCBI gene IDs |
| **Values** | Numeric (continuous); gene-level CN estimates |
| **Scale** | Log2 ratio relative to diploid reference (typically -2 to +3) |
| **Method** | SNP microarray or WES-derived CN calling |
| **Interpretation** | 0 = diploid (2 copies); <0 = deletion; >0 = amplification |


# Important Limitations & Disclaimers 

1. **Data Subset**: This package includes a curated subset for testing/examples. 
For comprehensive analyses, download the full DepMap Portal data.

2. **Licensing & Usage**: DepMap data is publicly available but has specific usage terms.
Verify compliance with your intended use: https://depmap.org/portal/documentation/

3. **Citation**: Always cite both DepMap (original source) and gDRtestData package.

# SessionInfo {-}

```{r sessionInfo}
sessionInfo()
```
