---
title: "gDRtestData"
author: "gDR team"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{gDRtestData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
library(gDRtestData)
```

# Overview

The `gDRtestData` package is intended to store and generate example data that can be used through the `gDR` suite.

# Use Cases

## Synthetic data generation

Since `gDR`is a computational suite for drug response data from any experiment, a synthetic dataset is also needed for testing and exploration purposes. 

The basis of everything are two functions to generate the synthetic sets of cell lines and drugs.

```{r}
cell_lines <- create_synthetic_cell_lines()
drugs <- create_synthetic_drugs()
```

In further step user may generate set of synthetic results:

1. Hill coefficient
```{r}
hill <- generate_hill_coef(cell_lines, drugs) 
head(hill)
```

2. EC50 metric
```{r}
ec50_met <- generate_ec50(cell_lines, drugs) 
```

3. E inf metric
```{r}
einf_met <- generate_e_inf(cell_lines, drugs)
```

The user can create synthetic raw data

```{r}
raw_data <- prepareData(create_synthetic_cell_lines(), create_synthetic_drugs())
```


## Synthetic object of gDR data model

The gDR data model is built on the MultiAssayExperiments (MAE) structure. Detailed description of gDR data model may be found in gDRcore package vignette. 

In `inst/testdata` user may find a set of `qs` file that are example of gDR data model for different data type. In file `synthetic_list.yml` maybe be found list of these dataset.

In `inst/scripts` you can find a script `generate_example_data.R` showing how to generate and process above mentioned data sets.
All key functions in this script can be found in `gDRcore/R/generate_wrappers.R`. 


# Package installation

The function `installAllDeps` assists in installing package dependencies.  

# SessionInfo {-}

```{r sessionInfo}
sessionInfo()
```
